### 2008-08-01 doc/backend_interface/Backends.tex
Initial version

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {C.U.Soettrup},
  pdftitle ={Arc backend infrastructure},
  pdfsubject = {Description and developer guide},
  pdfkeywords = {ARC,NorduGrid,manual,Backend,LRMS, Infosystem},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%

@@

{\raggedleft NORDUGRID-TECH-13\par}

@@

\begin{center}
\textit{\large Description and developer's guide}
\par\end{center}{\large \par}

@@

{\centering \large Adrian Taga\footnote{v.a.taga@fys.uio.no}, Thomas Frågåt\footnote{thomasf@fys.uio.no}, C.U.Søttrup, Bal\'azs K\'onya, G\'abor R\H oczei \par}

@@

This document describes the Advanced Resource Connector\cite{ARC}
batch system back-end infrastrucure. It aims to describe the infrastructure
in enough detail that a developer can add a new local resource management
system (LRMS) to the ARC middleware. It also intends to serve as a
reference manual describing existing batch system interfaces. Note
that the current solution described in this manual is not a real interface,
it is too diffuse to be that.

The batch system back-ends are what tie the ARC grid middleware (through
the Grid Manager (GM\cite{GM}) and the Information System layer\cite{INFOSYS})
to the underlying cluster management system or LRMS. The back-ends
consist of set of shell and Perl scripts whose role are twofold:

\begin{enumerate}
\item to allow the GM to manipulate jobs on the LRMS including job submit,
cancel operations and scanning for completed jobs
\item to collect information about jobs, users, batch system and the cluster
itself for the Information System. 
\end{enumerate}
The former will be referred to as the job control back-end interface
while the latter is the information system interface of the batch
system back-ends. These two will be treated separately in the following
sections.

\subsection{ARC Classic}
As of ARC version 0.6 the the two sets of scripts are located in different
directories within the NorduGrid subversion tree\cite{subversion}.
The job control interface scripts can be found under \texttt{arc\_development\_line/grid-manager/LRMS/}
while the information collectors are located at \texttt{arc\_development\_line/infosystem/.}

Section 2, 3, and 4 gives a description of the classic ARC back-ends scripts.

\subsection{ARC1}
As of ARC version 0.9, which is to be the next generation ARC, the scripts are located in different directories within the NorduGrid subversion tree\cite{subversion} as well. 
The architecture of ARC1 is different from Classic ARC so the job control interface scripts can be found under
\texttt{arc1/trunk/src/services/a-rex/lrms/}
while the information collectors are located in \texttt{arc1/trunk/src/services/a-rex/infoproviders/}.

Furthermore, the next generation of ARC supports both the classic NorduGrid information schema\cite{INFOSYS} and
currently a minimal set of the GLUE specification version 2.0 schema\cite{GLUE2}.

The ARC1 back-end architecture and GLUE2 information model implementation is presented in Section 5.

@@

Job control part of the LRMS interface is handled by the Grid Manager\cite{GM}.
It takes care of preparing a native batch system submission script,
managing the actual submission of the batch system job, cancellation
of job on request and scanning for completed batch jobs. Besides the
LRMS job control interface it is also the GM which provides e.g. the
data staging and communication with the grid client, provides RTE
environments, arranges file staging (to the node via LRMS capability),
dealing with stdout/stderr, etc. The job control batch system interface
of the GM requires three programs. These programs can be implemented
any way the designer sees fit, but all the existing back-end interfaces
use shell scripting for portability and ease of tailoring to a specific
site. The GM will call the following programs: cancel-LRMS-job, submit-LRMS-job,
and scan-LRMS-job. LRMS is replaced with the short hand name for the
LRMS e.g. cancel-pbs-job. The scripts are described one by one in
the following subsections. Useful information can also be found in
the Section \char`\"{}8.6 LRMS Support\char`\"{} and Section \char`\"{}8.7
Runtime Environment\char`\"{} of the Grid-Manager guide\cite{GM}.

@@

The main purpose of the information system batch interface is to populate
the nordugrid information model with locally collected information
obtained from the batch system. Important to recall that the locally
collected information is broader than what the batch-system can offer,
information taken from the grid-layer (mostly grid manager), from
the front-end machine and from the \texttt{arc.conf} configuration
file are also needed and used to populate the nordugrid information
model\cite{INFOSYS}. Below only the batch system related information
collection is covered. Note that the separation of the different sources
is not always that straightforward (e.g. the grid-manager information
has some overlap with the batch system info).

The information system interface consists of set of Perl scripts which
populate the local lDAP database by generating ldif files. Historically
two information provider \char`\"{}frameworks\char`\"{} have been
coexisted. As of ARC version 0.6 the old framework is being phased
out by the new framework. Briefly about the two frameworks:

\begin{itemize}
\item the old framework separated the non-batch system related functionality
into the \texttt{InfosysCluster.pm, InfosysQJU.pm} modules and kept
the batch system specific interface within the \texttt{cluster-condor.pl,
queue+jobs+users-condor.pl} (replace condor with your scheduler name)
files. There are two templates provided for batch system interface
developers: \texttt{cluster-gm.pl, queue+jobs+users-gm.pl}
\item the new framework originally proposed by Juha Lento%
\footnote{the original posting of the new framework proposal: http://staff.csc.fi/juha.lento/grid/rewrite.html%
} has been brought up-to-date by implementing all the features present
in the old framework. The new framework consists of two main scripts,
the \texttt{cluster.pl} and the \texttt{qju.pl}. The non-batch system
functionality is implemented by these script (plus some auxiliary
scripts). The batch system specific interface is moved to LRMS modules
such as \texttt{SGE.pm, PBS.pm, Fork.pm}, a template module \texttt{Template.pm}
is also provided. The main advantage of the new framework is the separation
of output generation from the information collection (this will ease
the migration to XML output).
\end{itemize}
The following subsection describes the LRMS interface part of the
information system as implemented via the new framework. To support
a particular LRMS a dedicated Perl module should be written that implements
the interface presented in the next subsections. Furthermore hooks
should be added to the \texttt{LRMS.pm} module, so that it uses the
correct Perl module depending on the LRMS type. Short description
of the files of the new framework:

\begin{itemize}
\item \texttt{cluster.pl and qju.pl}: main scripts called to populate the
nordugrid information model
\item \texttt{Shared.pm:} subroutines shared by cluster.pl and cluster.pl
\item \texttt{LogUtils.pm}: logging utilities
\item \texttt{LRMS.pm:} defines the LRMS interface (described in the next
subsection)
\item \texttt{Template.pm:} module template for implementing a new LRMS
interface
\item \texttt{SGE.pm, PBS.pm, LSF.pm, Fork.pm, LL.pm, Condor.pm}: actual
LRMS modules
\end{itemize}

@@

\begin{tabular}{|c|c|}
\hline 
key&
value\tabularnewline
\hline
\hline 
status&
LRMS jobstatus mapping%
\footnote{LRMS job states mapping is described in\cite{INFOSYS} %
}:Running->'R', Queued->'Q', Suspended->'S', Exiting->'E', Other->'O'\tabularnewline
\hline 
rank&
the job's position in the LRMS queue\tabularnewline
\hline 
mem&
the memory usage of the job in kBs\tabularnewline
\hline 
walltime&
consumed wall-time in minutes\tabularnewline
\hline 
cputime&
consumed CPU-time in minutes\tabularnewline
\hline 
reqwalltime&
wall-time requested by job in minutes\tabularnewline
\hline 
reqcputime&
CPU-time requested by job in minutes\tabularnewline
\hline 
nodes&
list of execution hosts\tabularnewline
\hline 
comment&
array of string comments about the job in LRMS, can be an empty array\tabularnewline
\hline
\end{tabular}

@@

\begin{tabular}{|c|c|}
\hline 
key&
value\tabularnewline
\hline
\hline 
freecpus&
number of freely available CPUs with their time limits in minutes
(see nordugrid-authuser-freecpus in\cite{INFOSYS})\tabularnewline
\hline 
queuelength&
estimated queue length for the specified user\tabularnewline
\hline
\end{tabular}

@@

The ARC configuration manual contains detailed information on all
the ARC configuration parameters\cite{config}and should be consulted
as primary source. Below the PBS specific variables are collected.

@@

\subsection{Condor}

The Condor \cite{Condor} system, developed at the University of Wisconsin-Madison,
was initially used to harness free cpu cycles of workstations. Over
time it has evolved into a complex system with many grid-oriented
features. Condor is available on a large variety of platforms.

Install Condor on the Grid Manager (GM) node and configure it as a
submit machine. Next, you must add the following to the node's Condor
configuration (\texttt{CONDOR\_IDS} can also be an environment variable):

\texttt{MAIL = <ARC\_install\_prefix>/libexec/finish-condor-job}

\texttt{CONDOR\_IDS = 0.0}

The \texttt{MAIL} attribute will instruct Condor to run the specified
program on job completion. The default on Condor is to run /bin/mail
to notify the user, but in this case, it is the GM that needs the
notification. Therefore, /bin/mail is replaced with a program especially
written for talking to the GM.

\texttt{CONDOR\_IDS} has to be 0.0, so that the above notification
program can access the Grid job's session directories (needed to extract
the job exit code from the Condor log).

Make sure that no normal users are allowed to submit Condor jobs from
this node. For one thing, it would not work for the user, since Condor
will try to notify the GM instead of the job owner on job completion.
If you don't allow normal user logins on the GM machine, then you
don't have to do anything. If you for some reason want to allow users
to log into the GM machine, simply don't allow them to execute the
condor\_submit program. This can be done by putting all local Unix
users allocated to the Grid in a single group, e.g. 'griduser', and
then setting the file ownership and permissions on condor\_submit
like this:

\texttt{chgrp griduser \$condor\_location/bin/condor\_submit}

\texttt{chmod 750 \$condor\_location/bin/condor\_submit }

\begin{itemize}
\item The Condor batch system back-end is enabled by setting \texttt{lrms=\char`\"{}condor\char`\"{}}
in the \texttt{{[}common]} configuration block. 
\item \texttt{condor\_location} configuration variable of the \texttt{{[}common]}
block should be set to the Condor install prefix (i.e., the directory
containing Condor's bin, sbin, etc).
\item \texttt{condor\_config} configuration variable of the \texttt{{[}common]}
block should be set to the value the environment variable \texttt{CONDOR\_CONFIG}
should have (but don't try to use the environment variable directly
as \texttt{\$CONDOR\_CONFIG}, since it will probably not be defined
when arc.conf is parsed!)
\item \texttt{condor\_rank} configuration variable of the \texttt{{[}common]}
block, if defined, will cause the \texttt{Rank} attribute to be set
in each job description submitted to Condor. Use this option if you
are not happy with the way Condor picks out nodes when running jobs
and want to define your own ranking algorithm. \texttt{condor\_rank}
should be set to a ClassAd float expression that you could use in
the \texttt{Rank} attribute in a Condor job description. For example:
\end{itemize}
\begin{quote}
\texttt{condor\_rank=\char`\"{}(1-LoadAvg/2){*}(1-LoadAvg/2){*}Memory/1000{*}KFlops/1000000\char`\"{} }
\end{quote}
\begin{itemize}
\item \texttt{condor\_requirements} configuration variable of the \texttt{{[}queue]}
block defined a subpool of condor nodes. See next section for details.
\end{itemize}

The job control part of the interface uses the \texttt{condor\_submit}
command to submit jobs. Some of the options used in the job's ClassAd
are:

\texttt{Requirements} -- is used to select the nodes that may run
the job. This is how ARC queues are implemented for Condor.

\texttt{Periodic\_remove} -- is used to enforce cputime and walltime
limits.

\texttt{Log} -- the job's condor log file is parsed by the information
scripts to find out whether the job was suspended.

The information system component uses the following Condor commands:

\texttt{condor\_status -long} -- for collecting information about
nodes

\texttt{condor\_status -format \char`\"{}\%s\textbackslash{}n\char`\"{}
Machine -constraint '...'} -- for listing nodes that make up an ARC
queue.

\texttt{condor\_q -long -global} -- for monitoring running jobs.

\texttt{condor\_history -l jobid} -- for collecting information about
finished jobs. Further cues are taken from the job's condor log file
and the body of the email sent by Condor when a job completes.

\subsubsection{Configuring Queues}

Condor does not support queues in the classical sense. It is possible,
however, to divide the Condor pool in several sub-pools. An ARC 'queue'
is then nothing more than a subset of nodes from the Condor pool.

Which nodes go into which queue is defined using the \texttt{condor\_requirements}
configuration option in the corresponding \texttt{{[}queue]} section.
It's value must be a well-formed constraint string that is accepted
by a \texttt{condor\_status -constraint '...'} command. Internally,
this constraint string is used to determine the list of nodes belonging
to a queue. This string can get quite long, so, for readability reasons
it is allowed to split it up into pieces by using multiple \texttt{condor\_requirements}
options. The full constrains string will be reconstructed by concatenating
all pieces.

Queues should be defined in such a way that their nodes all match
the information available in ARC about the queue. A good start is
for the \texttt{condor\_requirements} attribute to contain restrictions
on the following: \texttt{Opsys}, \texttt{Arch}, \texttt{Memory} and
\texttt{Disk}. If you wish to configure more than one queue, it's
good to have queues defined in such a way that they do not overlap.
In the following example disjoint memory ranges are used to ensure
this:

\begin{quote}
\texttt{{[}queue/large] condor\_requirements=\char`\"{}(Opsys == \char`\"{}linux\char`\"{}
\&\& (Arch == \char`\"{}intel\char`\"{} || Arch == \char`\"{}x86\_64\char`\"{})\char`\"{}
condor\_requirements=\char`\"{} \&\& (Disk > 30000000 \&\& Memory
> 2000)\char`\"{}}

\texttt{{[}queue/small] condor\_requirements=\char`\"{}(Opsys == \char`\"{}linux\char`\"{}
\&\& (Arch == \char`\"{}intel\char`\"{} || Arch == \char`\"{}x86\_64\char`\"{})\char`\"{}
condor\_requirements=\char`\"{} \&\& (Disk > 30000000 \&\& Memory
<= 2000 \&\& Memory > 1000)\char`\"{}}
\end{quote}
Note that '\texttt{nodememory}' attribute in arc.conf means the maximum
memory available for jobs, while the Memory attribute in Condor is
the physical memory of the machine. To avoid swapping (and these are
probably not dedicated machines!), make sure that '\texttt{nodememory}'
is smaller than the minimum physical memory of the machines in that
queue. If for example the smallest node in a queue has 1Gb memory,
then it would be sensible to use \texttt{nodememory=\char`\"{}850\char`\"{}}
for the maximum job size.

In case you want more precise control over which nodes are available
for grid jobs, using pre-defined ClassAds attributes (like in the
example above) might not be sufficient. Fortunately, it's possible
to mark nodes by using some custom attribute, say \texttt{NORDUGRID\_RESOURCE}.
This is accomplished by adding a parameter to the node's local Condor
configuration file, and then adding that parameter to \texttt{STARTD\_EXPRS}:

\begin{quote}
\texttt{NORDUGRID\_RESOURCE = True }

\texttt{STARTD\_EXPRS = NORDUGRID\_RESOURCE, \$(STARTD\_EXPRS)}
\end{quote}
Now queues can be restricted to contain only 'good' nodes. Just add
to each \texttt{{[}queue]} section in arc.conf:

\begin{quote}
\texttt{condor\_requirements=\char`\"{} \&\& NORDUGRID\_RESOURCE\char`\"{} }
\end{quote}

Only Vanilla universe is supported. MPI universe (for multi-CPU jobs)
is not supported. Neither is Java universe (for running Java executables).
ARC can only send jobs to Linux machines in the Condor pool, therefore
excluding other unixes and Windows destinations. The session directory
must be on a network shared directory, visible from all worker nodes.

\subsection{LoadLeveler}

LoadLeveler(LL), or Tivoli Workload Scheduler LoadLeveler in full,
is a parallel job scheduling system developed by IBM.

The back-end should work fine with a standard installation of LoadLeveler.
For the back-end to report the correct memory usage and cputime spent,
while running. LoadLeveler has to be set-up to show this data in the
llq command. Normally this is turned off for performance reasons.
It is up to the cluster administrator to decide whether or not to
publish this information. The back-end will work whether or not this
is turned on.

Only the two basic LRMS config options are relevant for LoadLeveler:

\begin{itemize}
\item The LoadLeveler batch system back-end is enabled by setting \texttt{lrms=\char`\"{}ll\char`\"{}}
in the \texttt{{[}common]} configuration block. 
\item \texttt{ll\_bin\_path} configuration variable of the \texttt{{[}common]}
block must be set to the path of the LoadLeveler binaries.
\end{itemize}

The LoadLeveler back-end uses LoadLeveler's command line interface(CLI)
commands to submit and cancel jobs. All information in the information
system is similarly parsed from the output of CLI commands. It does
not parse any log files, nor does it use the binary APIs. The reason
that the back-end is completely based on the CLI is that the log files
are normally kept on another machine than the front end and that the
binary API for LL changes quite often. Often with each new version
of LL.

There is at the moment no support for parallel jobs on the LoadLeveler
back-end.

@@

\subsection{LSF}

Set up one or more LSF queues dedicated for access by grid users.
All nodes in these queues should have a resource type which corresponds
to the one of the the frontend and which is reported to the outside.
The resource type needs to be set properly in the 'lsb.queues' configuration
file. Be aware that LSF distinguishes between 32 and 64 bit for Linux.
For a homogeneous cluster, the 'type==any' option is convenient alternative.

Example: In lsb.queues set either of the following:

\begin{itemize}
\item RES\_REQ = type==X86\_64 
\item RES\_REQ = type==any 
\end{itemize}
See the '-R' option of the bsub command man page for more explanation.

The LSF back-end requires that the following options are specified:

\begin{itemize}
\item \texttt{lrms=\char`\"{}lsf\char`\"{}} in the \texttt{{[}common]} configuration
block. 
\item \texttt{lsf\_bin\_path} configuration variable of the \texttt{{[}common]}
block must be set to the path of the LSF binaries.
\item lsf\_profile\_path must be set to the filename of the LSF profile
that the back-end should use. 
\end{itemize}
Furthermore it is very important to specify the correct architecture
for a given queue in arc.conf. Because the architecture flag is rarely
set in the xRSL file the LSF back-end will automatically set the architecture
to match the chosen queue. LSF's standard behaviour is to assume the
same architecture as the frontend. This will fail for instance if
the frontend is a 32 bit machine and all the cluster resources are
64 bit.If this is not done the result will be jobs eing rejected by
LSF because LSF believes there are no useful resources available.

The LSF implementation of the back-end are based solely on parsing
and running LSF's command line interface commands. No log files or
other methods are used. To get the correct output o any output at
all the back-end needs to have an apropriate LSF profile. The path
to this profile must be set in arc.conf. It will then be executed
by the back-end before running any of LSF's CLI commands.

Parallel jobs have not been tested on the LSF back-end.

The back-end does not at present support reporting different number
of free CPUs per user.

@@

\section{ARC1}
This section contains the current implementation of the ARC1 back-ends.

\subsection{The Information System}
ARC1 provides support for both the classic NorduGrid information schema\cite{INFOSYS} as well as an early minimal GLUE 2 implementation. The functionality is divided between several information collecting scripts and the output is rendered in XML format. Further details about each script are given in the subsections below.
The current GLUE 2 implementation is given in section Section 5.2.

\subsubsection{The information Collector Scripts}
The information collector scripts are divided between several seperate scripts:
\begin{itemize}
\item \texttt{cluster+qju.pl} - driver for information collection. It calls all other infomation collectors and prints results in XML. The information collection is done by one single invocation of this script
\item \texttt{InfoCollector.pm} - base class for all information collectors (i.e.: all files \texttt{*Info.pm})
\item \texttt{InfoChecker.pm} - used by InfoCollector to validate options and results against a simple 'schema' (not XML Schema)
\item \texttt{GMJobsInfo.pm} - collects information about jobs from grid manager status files
\item \texttt{HostInfo.pm} - collects other information that can be collected on the front end (hostname, software version, disk space for users, installed certificates, Runtime environments ...)
\item \texttt{LRMSInfo.pm} - collects information that is LRMS specific (queues, jobs, local user limits ...)
\item \texttt{BATCH\_SYSTEM\_NAME.pm} - plugins for LRMSInfo. Only Fork, SGE and PBS are updated to the new framework
\item \texttt{ARC0ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the classic NG schema
\item \texttt{ARC1ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the GLUE2 schema
\item \texttt{ARC0ClusterSchema.pm} - description of the info produced by ARC0ClusterInfo
\item \texttt{ARC1ClusterSchema.pm} - description of the info produced by ARC1ClusterInfo
\end{itemize}

All information about the hosting environment and queues is collected by one invocation of a perl script called \texttt{cluster+qju.pl}. The XML output is printed to \texttt{STDOUT} and it includes two representations of the output data, namely the classic NorduGrid information schema and an incomplete GLUE2 schema representation.

\subsection{The ARC GLUE 2 implementation}
Currently, the the following GLUE2 classes are included in ARC1:
\begin{itemize}
\item ComputingService
\item ComputingEndpoint
\item ComputingShare
\end{itemize}

While the following classes are still missing (TODO: According to our table in wiki, there should be some other sections as well):
\begin{itemize}
\item ApplicationEnvironment
\item ExecutionEnvironment
\item ComputingActivity
\item Policy including MappingPolicy / AccessPolicy
\end{itemize}

The different classes with their attributes are described in their respective sections below.

\subsubsection{ApplicationEnvironment}

@@

\begin{thebibliography}{1}
\bibitem{ARC}\char`\"{}Advanced Resource Connector middleware for
lightweight computational Grids\char`\"{}. M.Ellert et al., Future
Generation Computer Systems 23 (2007) 219-240.

\bibitem{GM}The NorduGrid Grid Manager And GridFTP Server: Description
And Administrator's Manual {[}NORDUGRID-TECH-2], A.Konstantinov

\bibitem{INFOSYS}\char`\"{}The NorduGrid/ARC Information System\char`\"{}
{[}NORDUGRID-TECH-4], B.Kónya

\bibitem{subversion}NorduGrid subversion: http://svn.nordugrid.org/trac/nordugrid/browser

\bibitem{config}ARC configuration document: currently the \texttt{arc.conf.template}
file.

\bibitem{Condor}http://www.cs.wisc.edu/condor/

\bibitem{GLUE2} TODO: ADD REFERENCE
\end{thebibliography}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {T.Frågåt},
  pdftitle ={Arc batch system back-end interface guide with support for GLUE 2},
  pdfsubject = {Description and developer's guide for ARC1},
  pdfkeywords = {ARC,ARC1,NorduGrid,manual,Backend,LRMS,Infosystem,GLUE2},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%

@@

{\raggedleft NORDUGRID-TECH-XX\par}

@@

\begin{center}
\textit{\large Description and developer's guide for ARC1}
\par\end{center}{\large \par}

@@

{\centering \large Adrian Taga\footnote{v.a.taga@fys.uio.no}, Thomas Frågåt\footnote{thomasf@fys.uio.no} \par}

@@

This document describes the Advanced Resource Connector~\cite{arc1}
batch system back-end infrastrucure. It aims to describe the infrastructure
in enough detail that a developer can add a new local resource management
system (LRMS) to the ARC middleware. It also intends to serve as a
reference manual describing existing batch system interfaces. Note
that the current solution described in this manual is not a real interface,
it is too diffuse to be that.

The batch system back-ends are what tie the ARC grid middleware (through
(ARC Resource-coupled EXecution servcie, A-REX~\cite{arex}) to the underlying
cluster management system or LRMS. The back-ends consist of set of shell and
Perl scripts whose role are twofold:

\begin{enumerate}
\item to allow the GM, which is resided in A-REX, to manipulate jobs on the LRMS including job submit,
cancel operations etc.
\item to collect information about jobs, users, batch system and the cluster
itself for the Information System. 
\end{enumerate}
The former will be referred to as the job control back-end interface
while the latter is the information system interface of the batch
system back-ends. These two will be treated separately in the following
sections.

\subsection{ARC Classic}
The ARC Classic has its own solution, please refer to ~\cite{backends-arc0}.

\subsection{ARC1}
As of ARC version 0.9, which is to be the next generation ARC, the scripts are located in different directories
within the NorduGrid subversion tree~\cite{svn}. 
The job control interface scripts can be found under \texttt{arc1/trunk/src/services/a-rex/lrms/}
while the information collectors are located in \texttt{arc1/trunk/src/services/a-rex/infoproviders/}.

The next generation of ARC supports both the classic NorduGrid information schema~\cite{is} and
currently a minimal set of the GLUE specification version 2.0 schema~\cite{glue2}.

@@

Job control part of the LRMS interface is handled by the Grid Manager~\cite{gm}.
It takes care of preparing a native batch system submission script,
managing the actual submission of the batch system job, cancellation
of job on request and scanning for completed batch jobs. Besides the
LRMS job control interface it is also the GM which provides e.g. the
data staging and communication with the grid client, provides RTE
environments, arranges file staging (to the node via LRMS capability),
dealing with stdout/stderr, etc. The job control batch system interface
of the GM requires three programs. These programs can be implemented
any way the designer sees fit, but all the existing back-end interfaces
use shell scripting for portability and ease of tailoring to a specific
site. The GM will call the following programs: cancel-LRMS-job, submit-LRMS-job,
and scan-LRMS-job. LRMS is replaced with the short hand name for the
LRMS e.g. cancel-pbs-job. The scripts are described one by one in
the following subsections. Useful information can also be found in
the Section \char`\"{}8.6 LRMS Support\char`\"{} and Section \char`\"{}8.7
Runtime Environment\char`\"{} of the Grid-Manager guide~\cite{gm}.

@@

The main purpose of the information system batch interface is to populate
the nordugrid information model and the GLUE2 model with locally collected
information obtained from the batch system. Important to recall that the
locally collected information is broader than what the batch-system can
offer, information taken from the grid-layer (mostly grid manager), from
the front-end machine and from the \texttt{arc.conf} configuration
file are also needed and used to populate the nordugrid information
model~\cite{is} and the GLUE2 schema~\cite{glue2}.

The information system interface consists of set of Perl scripts which
generates an XML output to \texttt{STDOUT}.

The following subsection describes the LRMS interface part of the
information system. To support a particular LRMS a dedicated Perl module
should be written that implements the interface presented in the next
subsections. Furthermore hooks should be added to the \texttt{LRMS.pm} module,
so that it uses the correct Perl module depending on the LRMS type.

The information collector scripts are divided between several seperate scripts:
\begin{itemize}
\item \texttt{cluster+qju.pl} - driver for information collection. It calls all other infomation collectors and prints results in XML. The information collection is done by one single invocation of this script
\item \texttt{InfoCollector.pm} - base class for all information collectors (i.e.: all files \texttt{*Info.pm})
\item \texttt{InfoChecker.pm} - used by InfoCollector to validate options and results against a simple 'schema' (not XML Schema)
\item \texttt{GMJobsInfo.pm} - collects information about jobs from grid manager status files
\item \texttt{HostInfo.pm} - collects other information that can be collected on the front end (hostname, software version, disk space for users, installed certificates, Runtime environments ...)
\item \texttt{LRMSInfo.pm} - collects information that is LRMS specific (queues, jobs, local user limits ...)
\item \texttt{BATCH\_SYSTEM\_NAME.pm} - plugins for LRMSInfo. Only Fork, SGE and PBS are updated to the new framework
\item \texttt{ARC0ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the classic NG schema
\item \texttt{ARC1ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the GLUE2 schema
\item \texttt{ARC0ClusterSchema.pm} - description of the info produced by ARC0ClusterInfo
\item \texttt{ARC1ClusterSchema.pm} - description of the info produced by ARC1ClusterInfo
\end{itemize}

All information about the hosting environment and queues is collected by one invocation of the perl script named
\texttt{cluster+qju.pl}. The XML output is printed to \texttt{STDOUT} and it includes
two representations of the output data, namely the classic NorduGrid information schema and an incomplete GLUE2 schema representation.

@@

\begin{tabular}{|c|c|}
\hline 
key&
value\tabularnewline
\hline
\hline 
status&
LRMS jobstatus mapping%
\footnote{LRMS job states mapping is described in~\cite{is} %
}:Running->'R', Queued->'Q', Suspended->'S', Exiting->'E', Other->'O'\tabularnewline
\hline 
rank&
the job's position in the LRMS queue\tabularnewline
\hline 
mem&
the memory usage of the job in kBs\tabularnewline
\hline 
walltime&
consumed wall-time in minutes\tabularnewline
\hline 
cputime&
consumed CPU-time in minutes\tabularnewline
\hline 
reqwalltime&
wall-time requested by job in minutes\tabularnewline
\hline 
reqcputime&
CPU-time requested by job in minutes\tabularnewline
\hline 
nodes&
list of execution hosts\tabularnewline
\hline 
comment&
array of string comments about the job in LRMS, can be an empty array\tabularnewline
\hline
\end{tabular}

@@

\begin{tabular}{|c|c|}
\hline 
key&
value\tabularnewline
\hline
\hline 
freecpus&
number of freely available CPUs with their time limits in minutes
(see nordugrid-authuser-freecpus in~\cite{is})\tabularnewline
\hline 
queuelength&
estimated queue length for the specified user\tabularnewline
\hline
\end{tabular}

@@

Below the PBS specific configuration variables are collected.

@@

@@

@@

\section{The ARC GLUE 2 implementation}
ARC1 provides support for both the classic NorduGrid information
schema~\cite{is} as well as an early minimal GLUE 2 implementation.
The output is rendered in XML format.

Currently, the the following GLUE2 entities are included in ARC1:
\begin{itemize}
\item ComputingService
\item ComputingEndpoint
\item ComputingShare
\end{itemize}

While the following entities are still missing (TODO: According to our table in wiki, there should be some other sections as well):
\begin{itemize}
\item ApplicationEnvironment
\item ExecutionEnvironment
\item ComputingActivity
\item Policy including MappingPolicy / AccessPolicy
\end{itemize}

The different entities with their attributes are described in their respective sections below.

\subsection{ComputingService}

@@

\subsection{ComputingEndpoint}

\subsection{ComputingShare}

\bibliography{grid}
